Learning PyEMMA : v\versionnumber

### 2018-07-05 manuscript.tex
added software section, fixed typos

\newcommand{\versionnumber}{0.1}
\newcommand{\githubrepository}{\url{github.com/markovmodel/pyemma_tutorials}}

@@

\author[1\authfn{1}*]{AUTH0}
\author[1\authfn{1}*]{AUTH1}
\author[1*]{AUTH2}
\affil[1]{AFFILIATION}

\corr{auth0@email}{A0}
\corr{auth1@email}{A1}
\corr{auth2@email}{A2}

@@

PyEMMA~\cite{pyemma} (http://emma-project.org) is a framework for the analysis of molecular dynamics (MD) simulations using Markov state models; the package is written in Python (\url{http://python.org}) with heavy usage of numpy~\cite{numpy} and compatible to scikit-learn~\cite{sklearn} package.

@@

In this tutorial, we assume familiarity with the theory behind the MSM approach and focus solely on usage of PyEMMA. In the first four sessions/notebooks, we introduce the basic components of the PyEMMA analysis pipeline, i.e., data I/O and featurization, dimension reduction and discretization; MSM estimation and validation, and MSM coase graining and analysis. In three further sessions/notebooks we introduce special topics, i.e., use of hidden MSMs, working with expectations and observables, and finding good features using the VAMP score approach. In the eigths and last session/notebook, we show common problems with data and how PyEMMA reacts to these specific situations.

@@

@@

This tutorial guides you through the elementary pipeline of buildiung, estimating, validating, and analysing MSMs (notebook 00). The individual steps are explained on simple systems in specialised notebooks (01-08).

@@

data-i/o + featurisation (number of features or system parameters)

dimension reduction and discretization (features plot, free energy or density fot tics[0,1]?)

@@

All notebooks are available on \githubrepository where you also might find additional materials and instalation instructions.

>>>>

\newcommand{\versionnumber}{0.1} % TODO: can we extract this number from a git tag?
\newcommand{\githubrepository}{\url{github.com/markovmodel/pyemma_tutorials}}

@@

\author[1\authfn{1}*]{AUTH0}
\author[1\authfn{1}*]{Martin K. Scherer}
\author[1*]{AUTH2}
\affil[1]{AFFILIATION}

\corr{auth0@email}{A0}
\corr{m.scherer@fu-berlin.de}{A1}
\corr{auth2@email}{A2}

@@

PyEMMA~\cite{pyemma} (http://emma-project.org) is a framework for the analysis of molecular dynamics (MD) simulations using Markov state models;
the package is written in Python (\url{http://python.org}) with heavy usage of NumPy~\cite{numpy} and compatible to scikit-learn~\cite{sklearn} package.

@@

In this tutorial, we assume familiarity with the theory behind the MSM approach and focus solely on usage of PyEMMA. 
In the first four sessions/notebooks, we introduce the basic components of the PyEMMA analysis pipeline, i.e., data I/O and featurization, dimension reduction and discretization; 
MSM estimation and validation, and MSM coarse graining and analysis. In three further sessions/notebooks we introduce special topics, i.e., 
use of hidden MSMs, working with expectations and observables, and finding good features using the VAMP score approach.
In the eigths and last session/notebook, we show common problems with data and how PyEMMA reacts to these specific situations.

@@

We utilize Jupyter notebooks to show code examples along with figures and interactive widgets to display molecules.
The user can install all needed packages with only one command, if he or she uses the conda command provided by the Anaconda Python stack.
We rely on Anaconda, because it resolves and installs dependencies and provides pre-compiled versions of hard to build packages like SciPy.
The installation will contain a launcher command to start the Jupyter notebook server
as well as the notebook files. The data for the demonstrated test systems is downloaded upon the first usage and is cached for future invocations of the tutorial.
The underlying software stack for running the tutorial importantly consists out of:
\begin{itemize}
 \item NumPy - number crunching
 \item SciPy - linear algebra, sparse computations
 \item Matplotlib - Visualization
 \item \textbf{PyEMMA} - MSM/HMM estimation, analysis, validation and visualization
 \item mdtraj - reading MD simulation data, calculate observables from MD
\end{itemize}

The software is currently supported for Python versions 3.5 and 3.6 on the operating systems Linux, OSX and Windows.

If one does not want to use Anaconda, he or she can also perform a manual installation via the pip installer or use the Binder to look and execute the tutorials online in any browser.

@@

This tutorial guides you through the elementary pipeline of building, estimating, validating, and analyzing MSMs (notebook 00). 
The individual steps are explained on simple systems in specialized notebooks (01-08).

@@

data-i/o + featurization (number of features or system parameters)

dimension reduction and discretization (features plot, free energy or density for tics[0,1]?)

@@

All notebooks are available on \githubrepository where you also might find additional materials and installation instructions.
